\chapter{Conclusion}
The course has been an interesting experience, allowing us to confirm our pre-existing assumptions as well as gaining new knowledge about ourselves. In the end we managed to create a functional prototype for Conductor Hero, an achievement we are all proud of. We believe this was due to our clear expectations, goals, and a common desire to participate. Good communication ensured that we had everything under control most of the time, and that we did not waste unnecessary time, maximizing work and limiting risk. Additionally, the group diversity and range of previous experiences helped us make choices that was beneficial to the project. The inclusive nature of the group helped ensure that everyone felt welcome, allowing them to present their ideas without fear of scrutiny. We believe all these factors contributed to the general success of our group, and taught us lessons that will be invaluable for the future.

\section{Applying teamwork skills in the future}
We believe that teamwork skills will be relevant for all team members in their future professional careers. While we all had some experience with teamwork before, there are a few aspects of teamwork we believe the course has made us more aware of.

First, we found the 6-animals model~\cite{six_animal_model} a very interesting method for group formation, and for ensuring a balance of diverse “traits” and “functions” within a group. We also believe that we have become more aware of how to take advantage of diversity within groups, while also staying aware of each group member’s needs and motivations. In this connection, we also believe that we have become more aware of the importance of clear and open communication, especially in cross-disciplinary environments. 

The importance of creating clear guidelines, in the forms of establishing group goals and expectations, is a tool we will utilize in future work. We believe a part of our success is due to establishing firm rules for the group at the start of the project. Contrasting this with previous group experiences without these rules, we see a significant difference in the outcomes. Establishing ground rules, and being clear on expectations, is something we will take with us into the future, and we believe it will enhance our practices.

While we experienced some minor situations, we did not experience any particularly difficult conflicts during the project. We do, however, believe that we are likely to encounter more challenging situations in the future. We have perhaps become more sensitive to when such situations are beginning to unfold, and have learnt some techniques that can be applied to deal with situations before they become destructive. 

In the end we are very thankful for the opportunities that has been presented to us through this course, and the valuable lessons we have learned.
